---
title: "EDS241: Final"
author: "Connor Flynn"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 
  
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
packages=c("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "tidyverse", "estimatr", "plm")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

#devtools::install_github('rstudio/rmarkdown')
options(scipen=999) # not scientific notation


```

```{r}
data <- read_csv(here("data/KM_EDS241.csv"))
```

(a) Using the data for 1981, estimate a simple OLS regression of real house values on the indicator for being located near the incinerator in 1981. What is the house value “penalty” for
houses located near the incinerator? Does this estimated coefficient correspond to the ‘causal’ effect of the incinerator (and the negative amenities that come with it) on housing values?
Explain why or why not.


```{r}
lm_robust(data = data, rprice ~ nearinc)

ggplot(data = data, aes(x = nearinc, 
                       y = rprice)) + 
  geom_smooth(method = "lm")
```
The house value “penalty” for houses located near the incinerator is $-24,456.64. This estimated coefficient does correspond to the 'causal' effect of the ininerator, and the negative ammenities that come with it, on housing values. 


(b) Using the data for 1978, provide some evidence the location choice of the incinerator was not “random”, but rather selected on the basis of house values and characteristics. [Hint: in the 1978 sample, are house values and characteristics balanced by nearinc status?]

```{r}
data_1978 <- data %>% 
  filter(year == 1978)
```

```{r}
data_1978_nearinc <- data_1978 %>% 
  filter(nearinc == 1)

mean_rprice_1978_nearinc <- mean(data_1978_nearinc$rprice)
median_rprice_1978_nearinc <- median(data_1978_nearinc$rprice)

mean_rprice_1978_nearinc
median_rprice_1978_nearinc
```


```{r}
data_1978_not_nearinc <- data_1978 %>% 
  filter(nearinc == 0)

mean_rprice_1978_not_nearinc <- mean(data_1978_not_nearinc$rprice)
median_rprice_1978_not_nearinc <- median(data_1978_not_nearinc$rprice)

mean_rprice_1978_not_nearinc
median_rprice_1978_not_nearinc

```

We can see the mean and median inflation-adjusted sales prices of houses in 1978 located near the incinerator are significantly lower than those located away from the incinerator. This suggests the placement of the incinerator was not random, and was purposely placed in an area with lower house values. 

(c) Based on the observed differences in (b), explain why the estimate in (a) is likely to be biased downward (i.e., overstate the negative effect of the incinerator on housing values).

Based on the observed differences in the mean and median of 1978 sales prices based on location, the linear regression in a is likely to overstate the the negative effect of the incinerator on housing values. Since housing prices in the area where the incineratoer was constructed were already significantly lower, the incinerator only exacterbated those differences and is not the sole reason for the difference in housing values. The incinerator could only be a causal effect if housing values in both areas were the same before it's construction, and all other variables were controlled for.

(d) Use a difference-in-differences (DD) estimator to estimate the causal effect of the incinerator on housing values without controlling for house and lot characteristics. Interpret the magnitude and sign of the estimated DD coefficient. 

```{r}
DD <- lm_robust(rprice ~ nearinc + as.factor(year),  data=data)
DD

se_DD <- starprep(DD, stat = c("std.error"), se_type = "HC2", alpha = 0.05)
se_DD
```

interpret

(e) Report the 95% confidence interval for the estimate of the causal effect on the incinerator in
(d).

answer

(f) How does your answer in (d) changes when you control for house and lot characteristics? Test the hypothesis that the coefficients on the house and lot characteristics are all jointly equal to 0.

```{r}
DD_control <- lm_robust(rprice ~ nearinc + age + rooms + area + land, data=data)
DD_control

se_DD_control <- starprep(DD_control, stat = c("std.error"), se_type = "HC2", alpha = 0.05)

se_DD_control
```

```{r}
se_models <- list(se_DD[[1]], se_DD_control[[1]])
stargazer(DD, DD_control, se = se_models, keep=c("nearinc"), type="text")
```



